\textbf{ Spectrophotometric parallaxes with \project{The Cygnet}:\\ Accurate distances for luminous red-giant stars }

### 2018-08-08 paper/sp.tex
notes to self
color cuts; differences in samples

% To-do items
% -----------
% - draft discussion
% - make sure we address the sensibleness and generality of the linear model for the photometric space
% - Coryn's point that we should carefully explain how our model is geometric.
% - related: HWR's point that this is standardizeable candles, like the Ia SNe results.
% - related: Put in an explanation of WHY this works so well.
% - related: Put in an explanation of how it is that we aren't really modeling luminosity or distance but really just generating parallaxes.
% - does the model need a cute name, like The Swan?
% - check with WISE experts that we have the correct WISE photometry
% - we need an uncertainty estimate or noise model for our outputs.
% - we have set lambda = 50 by x-val. We could go as high as 60 without hurting our results.
% - We probably should cut in color-color space. Probably in J-K vs B-R and in H-W2.
% - We will match axes on relevant plots.
% - ...

@@

% math macros
\newcommand{\T}{^{\mathsf{T}}}
\DeclareMathOperator*{\argmin}{argmin}
\newcommand{\logg}{\log g}
\newcommand{\BP}{B\!P}
\newcommand{\RP}{R\!P}
\newcommand{\gparallax}{\varpi^{(\mathrm{a})}}
\newcommand{\sparallax}{\varpi^{(\mathrm{sp})}}

@@

We need complete feature-vector information for all stars.  For this
reason, we define the Parent Sample to be the set of all stars that
meet the \apogee\ and \gaia\ cuts and also
have the complete set of photometry: $G_n$, $\BP_n$, $\RP_n$, $J_n$,
$H_n$, $K_n$, $W_{1n}$, and $W_{2n}$ and also an \apogee\ spectrum.
This Parent Sample contains ACE HOW MANY STARS? and is shown in \figurename~\ref{fig:parent}.
\begin{figure}
\includegraphics[width=\textwidth]{parent_sample.pdf}
\caption{A visualization of the Parent Sample stars. The left panel
  shows a color--color diagram for all XXXX stars in the combined Parent Sample, colored by
  the \apogee\ pipeline estimates of surface gravity. The right
  panel shows small spectral segments of four example stars,
  chosen to span metallicity and surface gravity.\label{fig:parent}}
\end{figure}

@@

In detail,
for each star $n$ in the Training Sets, we take from \gaia\ DR2
the astrometric parallax $\gparallax_n$ and its uncertainty $\sigma_n$.
Importantly, to each \gaia\ parallax measurement we add a positive
offset of $0.029$\,mas to adjust for the under-estimation of
parallaxes reported by the \gaia\ Collaboration (\citealt{lindegren}).
The Training Sets are shown in \figurename~\ref{fig:training}.
\begin{figure}
\includegraphics[width=\textwidth]{./training_sample.pdf}
\caption{A visualization of the Training Set stars. The left panel shows
  same color--color diagram as in \figurename~\ref{fig:parent}, but for
  the Training Set only. Note [HOGG SOME DIFFERENCES]. The right panel
  shows HOGG WHAT?\label{fig:training}}
\end{figure}

>>>>

% To-do items
% -----------
% - make sure we address the sensibleness and generality of the linear model for the photometric space
% - Coryn's point that we should carefully explain how our model is geometric.
% - related: HWR's point that this is standardizeable candles, like the Ia SNe results.
% - related: Put in an explanation of WHY this works so well.
% - related: Put in an explanation of how it is that we aren't really modeling luminosity or distance but really just generating parallaxes.
% - does the model need a cute name, like The Swan?
% - check with WISE experts that we have the correct WISE photometry
% - we need an uncertainty estimate or noise model for our outputs.
% - general HWR point that we aren't putting in to the model any idea about how apparent magnitude varies with distance; this means that we rely on data for EVERYTHING.
% - HWR will deliver cluster memberships, and then we will improve the cluster plots.
% - CODE: We will match axes on relevant plots.
% - CODE: we need to look at how things change with the parallax offset (0.029). It is another hyper-par.
% - CODE: we have set lambda = 50 by x-val -- final re-set?

@@

% math macros
\newcommand{\T}{^{\mathsf{T}}}
\DeclareMathOperator*{\argmin}{argmin}
\newcommand{\logg}{\log g}
\newcommand{\BP}{{G_\mathrm{BP}}}
\newcommand{\RP}{{G_\mathrm{RP}}}
\newcommand{\gparallax}{\varpi^{(\mathrm{a})}}
\newcommand{\sparallax}{\varpi^{(\mathrm{sp})}}

@@

We need complete feature-vector information for all stars.  For this
reason, we define the Parent Sample to be the set of all stars that
meet the \apogee\ and \gaia\ cuts and also
have the complete set of photometry: $G_n$, $\BP_n$, $\RP_n$, $J_n$,
$H_n$, $K_n$, $W_{1n}$, and $W_{2n}$ and also an \apogee\ spectrum.
In addition, we applied two light color cuts to remove stars with obviously
contaminated or outlying photometry:
\begin{eqnarray}
(J_n - K_n) &<& (+0.4\,\mathrm{mag}) + 0.45 * (\BP_n - \RP_n)
\\
(H_n - W_{2n}) &>& (-0.05\,\mathrm{mag})
\quad .
\end{eqnarray}
These cuts removed roughly 2~percent of the \apogee\ stars.
This Parent Sample contains 44808 (HOGG CHECK) and is shown in \figurename~\ref{fig:parent}.
\begin{figure}
\includegraphics[width=\textwidth]{parent_sample.pdf}
\caption{Color--color diagrams for the Parent Sample stars.
  The points are colored by
  the \apogee\ pipeline estimates of surface gravity.
  The color cuts used to trim outliers are shown as grey lines.\label{fig:parent}}
\end{figure}

@@

In detail,
for each star $n$ in the Training Sets, we take from \gaia\ DR2
the astrometric parallax $\gparallax_n$ and its uncertainty $\sigma_n$.
Importantly, to each \gaia\ parallax measurement we add a positive
offset of $0.029$\,mas to adjust for the under-estimation of
parallaxes reported by the \gaia\ Collaboration (\citealt{lindegren}).
The Training Sets are shown in \figurename~\ref{fig:training}.
\begin{figure}
\includegraphics[width=\textwidth]{./training_sample.pdf}
\caption{Color--color diagrams for the Training Set stars.
  Note that there is less color range in the Training Set than in the
  Parent Sample.\label{fig:training}}
\end{figure}

Comparison of {\figurename s}~\ref{fig:parent} and \ref{fig:training} shows
that there is more color range in the Parent Sample than in the Training Set.
This challenges the representativeness assumption in \sectionname~\ref{sec:assumptions}.
The principal reason for the difference is that the \gaia\ quality cuts
exclude stars preferentially from crowded regions, which also tend to be the
most dust-obscured.
The hope of the model assumptions is that the Training Set will contain sufficient
dust variation that the model will naturally learn the dust corrections and extrapolate
acceptably.
